---
title       : Week 4 presentation
subtitle    : Shiny app
author      : Nick Murray
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Synopsis

This presentation and the [accompanying Shiny app](https://flutable.shinyapps.io/ds09week4shiny/) shows air quality in Florey, a suburb to the north of Canberra. The author lives nearby. There are a high proportion of wood-burning fires in this part of the city, and I am interested in the 2.5μM and 10μM particulate readings.

I want to be able to "look back" a number of days and see the quantity of particulates I have breathed in - I'm a runner.

---
## Data 

Air quality data is from the [ACT government's data portal](http://www.health.act.gov.au/public-information/public-health/act-air-quality-monitoring/real-time-graphs). 

The downloadable dataset is ~12 MB, so to speed up the Shiny app, I preprocessed the data to extract relevant features, then uploaded to Dropbox.
The air quality data contains the following relevant variables.

- <b>PM10</b>       particulate matter less than 10 microns
- <b>PM2.5</b>      particulate matter less than 2.5 microns
- <b>Date</b>       date of reading
- <b>Time</b>       time of reading

```{r LoadAndPreProcChk,echo=FALSE}
options(scipen = 1, digits = 2)            # numbers > 10^5 in scientific notation, 2 sig figs
suppressPackageStartupMessages(library(zoo,       warn.conflicts = FALSE)) # for na.locf()
suppressPackageStartupMessages(library(dplyr,     warn.conflicts = FALSE))
suppressPackageStartupMessages(library(stringr,   warn.conflicts = FALSE))
suppressPackageStartupMessages(library(ggplot2,   warn.conflicts = FALSE))
suppressPackageStartupMessages(library(lubridate, warn.conflicts = FALSE))
suppressPackageStartupMessages(library(knitr,     warn.conflicts = FALSE))
suppressPackageStartupMessages(library(plotly,    warn.conflicts = FALSE))
suppressPackageStartupMessages(library(stringr,    warn.conflicts = FALSE))
# Download data if necessary
fnData <- "CanberraAirQuality.csv"

if ( !file.exists(fnData )) {
  #  message("Data file(s) missing, downloading....")
    download.file("https://www.data.act.gov.au/api/views/vm7b-spx8/rows.csv?accessType=DOWNLOAD&bom=true&format=true", fnData)
}
# Read data
if (!exists("airdata")) {
    #message("Air quality data missing, now reading. This may take a few seconds....")
    airdata <- read.csv(file = fnData, header = TRUE, sep = ",", stringsAsFactors = FALSE)
} 
```

```{r SubSetData,echo=FALSE}
# Fix invalid column name
colnames(airdata)[1] <- "Name"

# Remove Civic and Monash
airdata <- airdata[ airdata$Name %in% c("Florey"), ]
```

---
## Data processing 1
Data processing was mostly unremarkable, but here are some of the more interesting aspects of it:

Using %in% to select columns by name

```{r dp1,echo=TRUE}
# Keep only relevant columns
airdata <- airdata[ ,(colnames(airdata) %in% c("Name","PM10","PM2.5","Date","Time"))]
```

```{r dp2,echo=FALSE}
# Check for NAs in dates
datecheck <- sum(!is.na(strptime(airdata$Date,"%d-%b-%y"))) #should be 0 if all dates convert
if (datecheck > 0)
{stop("dates do not convert")}

# convert hour and day fields to a single POSIX format DateTime
airdata$DateTime <- str_c(airdata$Date, ' ', airdata$Time)
airdata$DateTime <- as.POSIXct(airdata$DateTime, format="%d %b %Y %H:%M:%S")

# remove Date and Time
airdata <- airdata[ , !(colnames(airdata) %in% c("Date", "Time"))]
```
Using the Zoo package's na.locf ("last observation carry forward") to impute NA values. See [StackOverflow](http://stackoverflow.com/questions/7735647/replacing-nas-with-latest-non-na-value) for more information.

```{r, dp3,echo=TRUE}
# impute missing values. If an NA, carry forward last observation
airdata$PM10  <- na.locf(airdata$PM10)
airdata$PM2.5 <- na.locf(airdata$PM2.5)
```

---
## Shiny app features
As Plotly's x/y axis control isn't reactive, I added a reactive x-axis ticks function to change the scaling:

         newticks <- reactive({
           if (input$lookbackdays <=5){  #lookbackdays is the duration you summarise over
             ticks <- 86400000.0/24      #There are 8.6 million milliseconds/day.
                                         #Ticks here are at hourly intervals
           }
            else{
              ticks <- 86400000.0/6      #For > 5 days, ticks are at 4-hourly intervals
            }      return(ticks)
         })

  
<!-- ---Remove these comments to embed Plotly graphs into the presentation   -->
<!-- ## Example  -->
<!-- ```{r} -->
<!-- # server side -->
<!-- # ensure date time data are in order, else you get some very odd graphs! -->
<!-- dt <- order(airdata$DateTime, decreasing = FALSE)  -->
<!-- airdata <- airdata[dt, ] -->
<!-- # define example lookback -->
<!-- lookbackdays <- 10 -->
<!-- lookback <- as.POSIXct(Sys.time() - (86400 * lookbackdays))  -->
<!-- # extract data from 'lookback' days ago -->
<!-- newairdata <- airdata[airdata$DateTime > lookback,  ]   #start from airdata so user doesn't have to reload it -->
<!-- newairdata$cumsumPM10   <- cumsum(newairdata$PM10) -->
<!-- newairdata$cumsumPM2.5  <- cumsum(newairdata$PM2.5) -->

<!--  pFlorey<- plot_ly(newairdata, x = ~DateTime, y = ~PM10, name = 'Particulates 10uM', type = 'scatter', mode = 'lines' , height = 200, width= 500) %>% -->
<!--             add_trace(y = ~ PM2.5, name='Particulates 2.5uM', mode='lines') %>% -->
<!--             layout(title = "Pollutants in Florey, last few days", -->
<!--               xaxis = list(title = "Date"), -->
<!--               yaxis = list (title = "Pollutant concentration")) -->

<!-- pFlorey -->
<!-- pFloreyCum <- plot_ly(newairdata, x = ~DateTime, y = ~ cumsumPM10, name="Cumulative sum of PM10 particulates",  -->
<!--                       type='scatter', mode='lines', fill='tozeroy', height = 200, width= 500) %>% -->
<!--            add_trace(y = ~ cumsumPM2.5, name="Cumulative sum of PM 2.5 particulates", fill='tozeroy') %>% -->
<!--             layout(title = str_c("Cumulative Pollutants in Florey, last ", lookbackdays, " days"), -->
<!--               xaxis = list(title = "Date"), -->
<!--               yaxis = list (title = "Pollutant concentration")) -->
<!-- pFloreyCum -->
<!-- ``` -->





